\hspace{1cm} Do syntactic and semantic similarity lead to interference effects? \newline Evidence from self-paced reading and event-related potentials using German

### 2024-08-26 paper/ResponseToReviewers.tex

\begin{quote}
    \Paste{future_work}
\end{quote}

\subsection*{Comment E.2}

\begin{quote}
``Second, a couple of Reviewer 1's comments (e.g., their point 2, and their final point about priors) relate to the fact that \textbf{there actually is some evidence for a degree of syntactic interference in aspects of the data, particularly perhaps in terms of the interaction and the combination with high semantic similarity, that you risk underplaying} (see also Reviewer 2's point 5).''
\end{quote}

\subsection*{Response to E.2}
We have re-done our analyses in response to the concerns raised by the reviewers and the Bayes factors from the revised analyses still do not provide evidence for syntactic interference or the interaction. As an overview of the revised Bayes factor analyses, we show Figures \ref{fig:spr_bfs} and \ref{fig:eeg_bfs} below. Figure \ref{fig:spr_bfs} shows the Bayes factor analysis of the self-paced reading data. There was evidence against syntactic interference under all priors in the pre-critical and critical region. Only in the spill-over region, there was anecdotal evidence (BF$_{10}$ = 2.8) for a small syntactic interference effect ($\pm$ 8 ms). There was evidence against the interaction under all priors in the pre-critical and spill-over regions. In the critical region, Bayes factors provided no evidence either way regarding the interaction (BF$_{10}$ = 1.4).

@@

\setcounter{figure}{4}
\Paste{fig_bf_spr}

Figure \ref{fig:eeg_bfs} shows the Bayes factor analyses of the ERP data in the spatio-temporal windows of the N400 and P600 (the latter is a new analysis which was added during the revision). For a very small syntactic interference effect on the N400 and a very small interaction on the P600 ($\pm$ 0.2\,$\mu V$), Bayes factors were inconclusive (BF$_{10}$ close to 1). For all other investigated effect sizes ($>$ 1\,$\mu V$) on both ERP components, Bayes factors provided evidence against syntactic interference and the interaction.

@@

\subsection*{Comment R1.8}
\begin{quote}
``Priors for Bayes factors\\
=============

Various priors are considered for Bayes factors. \textbf{Of course, it is good to explore several values but I wonder whether it is justified to conclude that syntactic interference or interaction should be excluded if they are not supported in models with unreasonably large priors (e.g., N+(0, 0.1)). Would anybody expect that interference is in range of 0.1 log-ms? As one of the authors of this manuscript nicely showed in Jager et al. (2017) and Schad et al. (2022), interference effects are small (at least for subject-verb agreement and ungrammatical illusions). Given that, I don't understand why we require a BF in favor of interference for a prior that is of size N+(0, 0.1). Shouldn't a smaller prior distribution suffice as evidence? This brings me again to my confusion in what the experiments actually show. The authors seem to argue that the only interference is the semantic one and they base the conclusion on large priors. But focusing on small priors (N+(0, 0.05) for spr, N-(0, 0.5) for EEG)), one could conclude that the interaction/syntactic interference might be present after all. Is there an independent reason to select the former priors over the latter for one's interpretation? To be sure, I don't want the authors to change the interpretation, but I would like to see either a more explicit argumentation why we should base our conclusion on such and such priors, or a more reflection in the text that explicitly states that the results are not equivocal and the current study is still compatible with a position that syntactic interference and/or interaction are real.}''\end{quote}

\subsubsection*{Response to R1.8}
The reviewer is right that syntactic interference and/or the interaction could represent real effects. However, our data does not provide conclusive evidence for them. Because it can be difficult to discern small differences in the Bayes factor plots (Figures 5 and 11 in Response to E.2), the Table below shows all Bayes factors for syntactic interference and the interaction. All Bayes factors above 1 are highlighted in bold. Following \textcite{jeffreys1998theory} and \textcite{lee2014bayesian}, we regard Bayes factors around 1 as providing no evidence either way (inconclusive). The only Bayes factor above 1 is BF10 = 2.8 for the syntactic interference effect in the reading times of the spill-over region. Because of the pre-critical effects in reading times, this effect in the spill-over region is potentially confounded: we cannot be sure whether the observed effect is due to syntactic retrieval interference or was somehow caused by semantic encoding interference. Therefore, we think that although the Bayes factor does provide anecdotal evidence for the syntactic interference manipulation in the spill-over region, we cannot conclude that the syntactic interference effect in our data is due to syntactic retrieval interference

@@

\end{table}

%BFs do not to need to be strongly in favor of an effect across all investigated priors to conclude that an effect is supported. However, to draw decisive conclusions in favor of an effect, the evidence for that effect should be stronger than anecdotal for at least a subset of reasonable priors.

Regarding the choice of priors, we believe that it is reasonable to explore a wide range of a priori effect sizes, specifically from $\pm$8 to $\pm$80 ms for the reading times. The meta-analysis of reading time studies by \textcite{jaeger_etal_2017} showed that the pooled estimate of inhibitory interference across studies with similar designs as ours was indeed small (13.1 ms, page 327), but some studies showed effects around 50 ms (e.g., two of the sub-experiments from \cite{vandyke07}) and one experiment by Franck et al., (2015) even showed an estimate above 100 ms (see Figure 2 on page 329 of \cite{jaeger_etal_2017}). Because such large interference effects have been found previously (see also our new Figure \ref{fig:previous_vs_present} showing all estimates of \cite{vandyke07} and \cite{mertzen}).

\newpage
We have revised the description of the SPR results to include more nuance regarding the BFs under the different priors:

### 2024-08-26 paper/main.tex
Updated via the Open Science Framework

Was there any evidence for the effects \textcolor{blue}{of interest }estimated above? The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}. The analyses use all the data from Experiments 1a and 1b. \Copy{SPR_bfs}{\textcolor{blue}{We found very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was weak (Normal(0, 0.01): BF$_{10}$ = 3.5,
Normal(0, 0.05): BF$_{10}$ = 1.3,
Normal(0, 0.1): BF$_{10}$ = 0.7). There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8). Similarly,} \Copy{insig_interaction}{\textcolor{blue}{there was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1). Only under the narrowest prior in the critical region, there was very weak evidence for the interaction (BF$_{10}$ = 1.4).}}\label{insig_interaction}

\textcolor{blue}{All in all, the Bayes factor analysis provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but at best weak evidence in the critical region. In contrast, the Bayes factor analysis provided at best weak evidence for syntactic interference and the interaction or even evidence against them.}}

@@

\textcolor{blue}{The Bayes factors under different priors on $F$ are shown in Figure \ref{fig:BF_compmods}. Compared to the null model, the one-cue model, the Bayes factors indicate strong evidence for the new two-cues model; this comparison shows that there is strong evidence for interference effects. Compared to the three-cues model proposed by \textcite{mertzen}, the Bayes factors provide either no evidence for the two-cues model or anecdotal to moderate evidence for the new two-cues model depending on the prior. This model comparison shows that there is some support for the idea that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the previously assumed \{$\pm$grammatical subject\} cue. This result holds for prior assumptions that allows higher probability mass for small values of $F$., i.e., for $F<0.01$.}

>>>>

Was there any evidence for the effects \textcolor{blue}{of interest }estimated above? The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}. The analyses use all the data from Experiments 1a and 1b. \Copy{SPR_bfs}{\textcolor{blue}{We found very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was weak (Normal(0, 0.01): BF$_{10}$ = 3.5,
Normal(0, 0.05): BF$_{10}$ = 1.3,
Normal(0, 0.1): BF$_{10}$ = 0.7). There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8).} \Copy{insig_interaction}{\textcolor{blue}{There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1). Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4).}}\label{insig_interaction}

\textcolor{blue}{All in all, the Bayes factor analysis provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but at best weak evidence in the critical region. In contrast, the Bayes factor analysis across all regions provided at best weak evidence for syntactic interference or even evidence against it and evidence against the interaction.}}

@@

\textcolor{blue}{The Bayes factors under different priors on $F$ are shown in Figure \ref{fig:BF_compmods}. Compared to the null model, the one-cue model, the Bayes factors indicate strong evidence for the new two-cues model; this comparison shows that there is strong evidence for interference effects. Compared to the three-cues model proposed by \textcite{mertzen}, the Bayes factors provide either no evidence for the two-cues model or anecdotal to moderate evidence for the new two-cues model depending on the prior. This model comparison shows that there is some support for the idea that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the previously assumed \{$\pm$grammatical subject\} cue. This result holds for prior assumptions that allows higher probability mass for small values of the scaling parameter $F$., i.e., for $F<0.01$.}
